---
title: "Final Data Story"
author: 
    - Marion Bauman
    - Matt Carswell
    - Thomas Sigall
    - Patricia Schenfeld
date: last-modified
date-format: long

execute:
  echo: false
engine: knitr
format:
    html:
        embed_resources: true
        toc: true
theme: morph
bibliography: ../../references.bib

# output: html_document
# format: 
#   html:
#     code-tools: true
#     embed-resources: true
#     grid:
#         body-width: 1100px
# engine: knitr
# theme: lux
# execute: 
#   echo: false
# vega:
#   renderer:
#     css: |
#       @import url("https://fonts.googleapis.com/css2?family=Nunito+Sans:wght@200;300;400;500;600;700&display=swap");
#       @font-face {
#         font-family: 'Nunito Sans';
#       }

#       .vega-embed .marks text {
#         font-family: 'Nunito Sans', sans-serif;
#       }
---

```{r}
# set conda environment
reticulate::use_condaenv("dsan5200", required = TRUE)
```

```{r}
color_palette <- c('#493FB3', '#D34B7F', '#FFBF49', '#8edce6', '#C894FF','#0B5351', '#4C230A')
#color_palette <- c("#A8C4FF", "#A56A6E", "#89B587", "#6A7FAB", "#FFDAA1", "#EFA4D6", "#7C5FBC", "#F2B37A")
```

# Introduction

Starting in 2022, a breakout social media trend started the rise of a troubling trend: teenagers stealing cars for fun. The trend, started by the alleged "Kia Boys," spread on the social media platform TikTok, where teenage boys posted videos of themselves stealing cars to take them on joyrides [@nyt-2023]. Often, these videos included step-by-step instructions on how to steal a car, and many of the cars stolen were Kia and Hyundai models. Videos show an easy trick that that involves just a screw driver and a USB cord to start the car in less than a minute due to a software error [@nyt-opinion]. In @fig-google-trends, we see a significant increase in Google searches for "Kia Boys", "Hyundai Theft", and "Kia Theft" starting in 2022 and continuing through today (2024).

{{< include content/google-search-trends.qmd >}}

# Who is Committing These Crimes

{{< include content/arrests-by-age.qmd >}}

{{< include content/difference-by-age.qmd >}}

# Where Are These Crimes Happening

## Percentage of Kia-Hyundai Thefts in the US by Month and Year (2020-2023)

{{< include content/kia_hyundai_pct_national.qmd >}}

## Car Thefts by Block in Washington, DC (2019-2023)

{{< include content/dc-thefts-leaflet.qmd >}}

## Car Thefts by Ward in Washington, DC (2019-2023)

{{< include content/observable_plot.qmd >}}

<!-- <iframe width="100%" height="800" frameborder="0"
  src="https://observablehq.com/embed/0de5730c9a255a03?cell=*"></iframe> -->

# When Are These Crimes Happening

## Total DC Car Theft Activity by Hour of the Day for Each Year (2019-2023)

{{< include content/car_theft_hour.qmd >}}

# Why Are These Crimes Happening

<!-- {{< include content/percent-change.qmd >}} -->
